format

### 2021-05-06 log_reg_analysis.Rmd
updated logistic regression analysis - no split (include all data), bootstrap x1000

@@

  theme(strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))
```

@@

  df_hbox_impute %>%
    filter(Status == status) %>% 
    dplyr::select(-c(subject_id, Status)) %>% 
    cor(method = "spearman")

@@

m_cor_hv <- df_hbox_impute %>% f_calc_cor(status = "HV")

m_cor_um <- df_hbox_impute %>% f_calc_cor(status = "UM")

m_cor_cm <- df_hbox_impute %>% f_calc_cor(status = "CM")

@@

      ggplot(aes(x = X1, y = X2, fill = value, label = round(value, 2))) + 
      geom_tile() +
      scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
      geom_text(size = 6) +

      xlab("") +
      ylab("") +
      theme(axis.text.x = element_text(angle = 45, hjust = 0.9),
            axis.text = element_text(size = 25),
            legend.text = element_text(size = 40),
            legend.title = element_text(size = 50),
            legend.key.size = unit(2, 'cm'),
            plot.title = element_text(size = 30)
            )

@@

- logistic regression model
- variables: DeoxyHb, DeoxyHb_alpha, OxyHb, OxyHb_alpha, O2sat, O2sat_alpha, THC, THC_alpha, THC_sd
- bootstrap x100
- BoxCox transformation

@@

# split data
set.seed(123)
hbox_split <- initial_split(df_hbox_voi_01, strata = Status)
hbox_train <- training(hbox_split)
hbox_test <- training(hbox_split)

# resample
set.seed(234)
hbox_boot <- bootstraps(hbox_train, times = 100)

@@

# recipe
hbox_rec <- recipe(Status ~ ., data = hbox_train) %>% 
  step_BoxCox(all_predictors()) %>% 
  step_nzv(all_predictors())

@@

```{r train_model}

@@

@@

Training data

@@

  ggtitle("Confusion matrix: training data") +
  theme_bw()

@@

```{r roc}

@@

Testing data
```{r test}

# fit to test data
hbox_final <- hbox_wf %>% 
  add_model(glm_spec) %>% 
  last_fit(hbox_split)

# compute metrics on testing data
collect_metrics(hbox_final) %>% 
  dplyr::select(-c(.estimator, .config)) %>% 
  as.data.frame()

# predictions
collect_predictions(hbox_final) %>% 
  count(.pred_class, Status) %>% 
  dplyr::rename("Prediction" = ".pred_class", "Truth" = "Status", "Freq" = "n") %>%  
  mutate(Prediction = ifelse(Prediction == 1, "CM", "UM")) %>% 
  mutate(Truth = ifelse(Truth == 1, "CM", "UM")) %>%

  ggtitle("Confusion matrix: testing data") +
  theme_bw()

@@

hbox_final$.workflow[[1]] %>% 
  tidy() %>% 
  as.data.frame()

@@

# OR
hbox_final$.workflow[[1]] %>% 
  tidy(exponentiate = TRUE) %>% 
  arrange(desc(estimate)) %>% 
  as.data.frame()

```{r plot_voi}

  ggplot(aes(x = THC_alpha, y = THC_sd, color = Status, fill = Status, size = O2sat_alpha)) +
  geom_point(alpha = 0.7) +
  stat_conf_ellipse(level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  stat_mean(shape = 1, na.rm = FALSE) +
  ggtitle("VOI") +
  theme_bw()

@@

- logistic regression model
- variables: THC_alpha, THC_sd (O2sat_alpha was NS when included in model, p = 0.42)
- bootstrap x100
- BoxCox transformation

@@

# split data
set.seed(456)
hbox_split2 <- initial_split(df_hbox_model2, strata = Status)
hbox_train2 <- training(hbox_split2)
hbox_test2 <- training(hbox_split2)

# resample
set.seed(789)
hbox_boot2 <- bootstraps(hbox_train2, times = 100)

@@

# recipe
hbox_rec2 <- recipe(Status ~ ., data = hbox_train2) %>% 
  step_BoxCox(all_predictors()) %>% 
  step_nzv(all_predictors())

@@

```{r train_model2}

glm_res2 <- hbox_wf2 %>% 
  add_model(glm_spec2) %>%
  fit_resamples(
    resamples = hbox_boot2,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = tune::control_resamples(save_pred = TRUE)
  )

@@

```{r roc2}

@@

Testing data
```{r test2}

# fit to test data
hbox_final2 <- hbox_wf2 %>% 
  add_model(glm_spec2) %>% 
  last_fit(hbox_split2)

# compute metrics on testing data
collect_metrics(hbox_final2) %>% 
  dplyr::select(-c(.estimator, .config)) %>% 
  as.data.frame()

# predictions
collect_predictions(hbox_final2) %>% 
  count(.pred_class, Status) %>% 
  dplyr::rename("Prediction" = ".pred_class", "Truth" = "Status", "Freq" = "n") %>%  
  mutate(Prediction = ifelse(Prediction == 1, "CM", "UM")) %>% 
  mutate(Truth = ifelse(Truth == 1, "CM", "UM")) %>%

  ggtitle("Confusion matrix: testing data") +
  theme_bw()

@@

hbox_final2$.workflow[[1]] %>% 
  tidy() %>% 
  as.data.frame()

@@

# OR
hbox_final2$.workflow[[1]] %>% 
  tidy(exponentiate = TRUE) %>% 
  arrange(desc(estimate)) %>% 
  as.data.frame()

@@

```{r plot_voi2}

df_hbox_model2 %>%

  ggplot(aes(x = THC_alpha, y = THC_sd, color = Status, fill = Status)) +
  geom_point(alpha = 0.7) +
  stat_conf_ellipse(level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  stat_mean(shape = 1, na.rm = FALSE) +
  ggtitle("VOI") +
  theme_bw()

@@

>>>>

load("glm_res.Rdata")

@@

  theme(strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

@@

  df %>%
    filter(Status == status) %>% 
    dplyr::select(-Status) %>% 
    cor(method = "spearman")

@@

m_cor_hv <- df_hbox_impute %>% dplyr::select(c(2, 11:19)) %>% f_calc_cor(status = "HV")

m_cor_um <- df_hbox_impute %>% dplyr::select(c(2, 11:19)) %>% f_calc_cor(status = "UM")

m_cor_cm <- df_hbox_impute %>% dplyr::select(c(2, 11:19)) %>% f_calc_cor(status = "CM")

@@

      ggplot(aes(x = X1, y = X2, fill = value, label = round(value, 2))) + 
      geom_tile() +
      scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
      geom_text(size = 8) +

      xlab("") +
      ylab("") +
      theme(axis.text.x = element_text(angle = 45, hjust = 0.9),
            axis.text = element_text(size = 35),
            legend.text = element_text(size = 40),
            legend.title = element_text(size = 50),
            legend.key.size = unit(2, 'cm'),
            plot.title = element_text(size = 30)
            )

@@

- logistic regression model
- variables: DeoxyHb, DeoxyHb_alpha, OxyHb, OxyHb_alpha, O2sat, O2sat_alpha, THC, THC_alpha, THC_sd
- bootstrap x1000
- BoxCox transformation

@@

# resample
set.seed(234)
hbox_boot <- bootstraps(df_hbox_voi_01, times = 1000)

@@

# recipe
hbox_rec <- recipe(Status ~ ., data = df_hbox_voi_01) %>% 
  step_BoxCox(all_predictors()) %>% 
  step_nzv(all_predictors())

@@

```{r train_model, eval = FALSE}

@@

```{r best}

# select best model to be used for evaluation
glm_best <- glm_res %>% select_best("roc_auc")

# finalize model
hbox_final <- hbox_wf %>% 
  add_model(glm_spec) %>% 
  finalize_workflow(glm_best) %>%
  fit(df_hbox_voi_01) %>%
  pull_workflow_fit()

@@

@@

  ggtitle("Confusion matrix") +
  theme_bw()

@@

```{r roc, eval = FALSE}

@@

@@

hbox_final %>% 
  tidy() %>% 
  as.data.frame()

@@

# OR
hbox_final %>% 
  tidy(exponentiate = TRUE) %>% 
  arrange(desc(estimate)) %>% 
  as.data.frame()

@@

- logistic regression model
- variables: THC_alpha, THC_sd
- bootstrap x1000
- BoxCox transformation

@@

# resample
set.seed(789)
hbox_boot2 <- bootstraps(df_hbox_model2, times = 1000)

@@

# recipe
hbox_rec2 <- recipe(Status ~ ., data = df_hbox_model2) %>% 
  step_BoxCox(all_predictors()) %>% 
  step_nzv(all_predictors())

@@

```{r train_model2, eval = FALSE}

glm_res2 <- hbox_wf2 %>% 
  add_model(glm_spec2) %>%
  fit_resamples(
    resamples = hbox_boot2,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
  )

save(glm_res, glm_res2, file = "glm_res.Rdata")

```{r best2}

# select best model to be used for evaluation
glm_best2 <- glm_res2 %>% select_best("roc_auc")

# finalize model
hbox_final2 <- hbox_wf2 %>% 
  add_model(glm_spec2) %>% 
  finalize_workflow(glm_best2) %>%
  fit(df_hbox_model2) %>%
  pull_workflow_fit()

@@

```{r roc2, eval = FALSE}

@@

@@

hbox_final2 %>% 
  tidy() %>% 
  as.data.frame()

@@

hbox_final2 %>% 
  tidy(exponentiate = TRUE) %>% 
  arrange(desc(estimate)) %>% 
  as.data.frame()

@@

## THC_alpha & THC_sd

```{r violin_plot}

df_hbox_log_voi %>% 
  pivot_longer(2:10, names_to = "variable", values_to = "value") %>% 
  filter(variable %in% c("THC_alpha", "THC_sd")) %>%

  ggtitle("Violin plot") +
  theme_bw() +
  theme(legend.position = "none")  +

  theme(strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 15))

@@

```{r box_plot}

# calculate quantiles of THC_alpha scores
Q1 <- quantile(df_hbox_log_voi$THC_alpha, 0.25)
Q2 <- quantile(df_hbox_log_voi$THC_alpha, 0.50)
Q3 <- quantile(df_hbox_log_voi$THC_alpha, 0.75)

df_hbox_log_voi %>% 
  dplyr::select(c("Status", "THC_alpha", "THC_sd")) %>%

  # assign quantiles to THC_alpha scores
  mutate(THCa_quartile = case_when(

    THC_alpha < Q1 ~ 1,
    Q1 < THC_alpha & THC_alpha < Q2 ~ 2,
    Q2 < THC_alpha & THC_alpha <= Q3 ~ 3,
    Q3 < THC_alpha ~ 4

  ) %>% as.factor()) %>%

  # plot
  ggplot(aes(x = Status, y = THC_sd)) +
  geom_boxplot(show.legend = FALSE, outlier.shape = NA) +
  geom_point(aes(color = THCa_quartile), position = position_jitter(0.2), size = 3) +
  stat_summary(aes(color = THCa_quartile), fun = "median", geom = "crossbar", size = 0.3, width = 0.5) +
  scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF", "black", "black")) +

  ggtitle("Relationship between THC_sd and THC_alpha") +
  theme_bw()

```{r scatter_plot}

cm_model <- lm(THC_sd ~ THC_alpha, data = df_hbox_log_voi %>% filter(Status == "CM"))
um_model <- lm(THC_sd ~ THC_alpha, data = df_hbox_log_voi %>% filter(Status == "UM"))

  ggplot(aes(x = THC_alpha, y = THC_sd, color = Status, fill = Status)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +

  geom_text(label = paste0("y = ", 
                           round(cm_model$coefficients[[1]], 2), 
                           " + ", 
                           round(cm_model$coefficients[[2]], 2), 
                           "x ; R^2 = ", 
                           round(summary(cm_model)$adj.r.squared, 2)),
            x = 1.3, y = 0.55,
            color = "#00BFC4",
            size = 6) +

    geom_text(label = paste0("y = ", 
                           round(um_model$coefficients[[1]], 2), 
                           " + ", 
                           round(um_model$coefficients[[2]], 2), 
                           "x ; R^2 = ", 
                           round(summary(um_model)$adj.r.squared, 2)),
            x = 1.3, y = 0.6,
            color = "#F8766D",
            size = 6) +

  ggtitle("Scatter plot") +
  theme_bw()
